Lineage Analysis of CRISPR and shRNA Screens

### 2018-09-18 lineages_18Q3.Rmd

For mutation calling, get paired gene name and cell line fields in a data frame. For this analysis, we don't care about how many mutations there are per gene or what type of mutations there are, so I didn't save more information. I took unique gene-cell line combinations since we only cared about mutation presence/absence. Add a `Mutation_Status` column denoting all entries in MAF files as mutations present in the associated cell lines.
```{r}
maf_raw <- read.delim("./data_munging/CCLE_DepMap_18q3_maf_20180718.txt.gz", header = TRUE, sep = "\t")
# Filter for cell lines in CRISPR screen
maf_raw <- filter(maf_raw, Broad_ID %in% unique(crispr_meta$Broad_ID))

# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "Tumor_Sample_Barcode", "Broad_ID", "Variant_Classification", "isDeleterious", "Reference_Allele"))
colnames(maf_df)[colnames(maf_df) == "Tumor_Sample_Barcode"] <- "CCLE_Name"
maf_df$Var_Length <- nchar(as.character(maf_df$Reference_Allele))

@@

```{r}
max(nchar(as.character(maf_raw$Reference_Allele)))
maf_raw %>% group_by(Variant_Type, isTCGAhotspot) %>% tally()

@@

```{r}
dep <- read.delim("./data_munging/gene_dependency_18Q3.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(dep) <- gsub(" .*", "", colnames(dep))
```

@@

crispr_ptmuts_lvlA_plot <- ggplot(data = crispr_ptmuts_lvlA, aes(x = Hugo_Symbol, y = Score)) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_violin(mapping = aes(fill = Mutation_Status_DeleteriousMissense), alpha = 0.7, position = position_dodge(0.85)) +
  geom_boxplot(mapping = aes(color = Mutation_Status_DeleteriousMissense), position = position_dodge(0.85), width = 0.1) +
  scale_fill_manual(values = crispr_ptmuts_lvlA_color) +
  scale_color_manual(values = c("black", "black")) +
  guides(color = FALSE) +
  theme(legend.direction = "horizontal", legend.justification = c(1, 0), legend.position = c(1, 0), legend.box.margin = margin(c(10, 10, 10, 10))) +
  labs(fill = "Mutation Status", y = "CERES Score", x = "Gene")
crispr_ptmuts_lvlA_plot
# ggsave("./plots_18Q3/manuscript/crispr_ptmuts_lvlA_plot.png", crispr_ptmuts_lvlA_plot, width = 12, height = 5, units = "in")
```

@@

```{r}
g2p_indications <- filter(read.delim("./data_munging/data_mutation_associations_appended.csv", sep = "\t", row.names = 1, header = TRUE), Evidence.Level == "A")

maf_g2p_indications <- filter(maf_raw, Genome_Change %in% g2p_indications$MutationName)

test <- merge(maf_g2p_indications, crispr_data, by = c("Hugo_Symbol", "Broad_ID"))
```

>>>>

For mutation calling, get paired gene name and cell line fields in a data frame. For this analysis, we don't care about how many mutations there are per gene or what type of mutations there are, so I didn't save more information. I took unique gene-cell line combinations since we only cared about mutation presence/absence. Add a `Mutation_Status` column denoting all entries in MAF files as mutations present in the associated cell lines.
```{r}
maf_raw <- read.delim("./data_munging/CCLE_DepMap_18q3_maf_20180718.txt.gz", header = TRUE, sep = "\t")
# Filter for cell lines in CRISPR screen
maf_raw <- filter(maf_raw, Broad_ID %in% unique(crispr_meta$Broad_ID))
colnames(maf_raw)[colnames(maf_raw) == "Tumor_Sample_Barcode"] <- "CCLE_Name"

# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "CCLE_Name", "Broad_ID", "Variant_Classification", "isDeleterious", "Reference_Allele"))
maf_df$Var_Length <- nchar(as.character(maf_df$Reference_Allele))

@@

```{r, eval=FALSE}
max(nchar(as.character(maf_raw$Reference_Allele)))
maf_raw %>% group_by(Variant_Type, isTCGAhotspot) %>% tally()

@@

```{r, eval=FALSE}
dep <- read.delim("./data_munging/gene_dependency_18Q3.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(dep) <- gsub(" .*", "", colnames(dep))
```

@@

crispr_ptmuts_lvlA_plot <- ggplot(data = crispr_ptmuts_lvlA, aes(x = Hugo_Symbol, y = Score)) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_violin(mapping = aes(fill = Mutation_Status_DeleteriousMissense), alpha = 0.7, position = position_dodge(0.85)) +
  geom_boxplot(mapping = aes(color = Mutation_Status_DeleteriousMissense), position = position_dodge(0.85), width = 0.1) +
  scale_fill_manual(values = crispr_ptmuts_lvlA_color) +
  scale_color_manual(values = c("black", "black")) +
  guides(color = FALSE) +
  theme(legend.direction = "horizontal", legend.justification = c(1, 0), legend.position = c(1, 0), legend.box.margin = margin(c(10, 10, 10, 10))) +
  labs(fill = "Mutation Status", y = "CERES Score", x = "Gene")
crispr_ptmuts_lvlA_plot
# ggsave("./plots_18Q3/manuscript/crispr_ptmuts_lvlA_plot.png", crispr_ptmuts_lvlA_plot, device = "png", width = 12, height = 5, units = "in")

crispr_ptmuts_lvlA_plot_faceted <- ggplot(data = crispr_ptmuts_lvlA, aes(x = Mutation_Status_DeleteriousMissense, y = Score)) +
  facet_wrap(~ Hugo_Symbol, ncol = length(unique(crispr_ptmuts_lvlA$Hugo_Symbol))) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_violin(mapping = aes(fill = Mutation_Status_DeleteriousMissense), alpha = 0.7, position = position_dodge(0.85)) +
  geom_boxplot(mapping = aes(color = Mutation_Status_DeleteriousMissense), position = position_dodge(0.85), width = 0.1) +
  scale_fill_manual(values = crispr_ptmuts_lvlA_color) +
  scale_color_manual(values = c("black", "black")) +
  guides(color = FALSE) +
  theme(legend.direction = "horizontal", legend.justification = c(0.5, 0), legend.position = c(0.5, 0), legend.box.margin = margin(c(10, 10, 10, 10)), axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  labs(fill = "Mutation Status", y = "CERES Score", x = "Mutation Status")
crispr_ptmuts_lvlA_plot_faceted
# ggsave("./plots_18Q3/manuscript/crispr_ptmuts_lvlA_plot_faceted.png", crispr_ptmuts_lvlA_plot_faceted, device = "png", width = 12, height = 5, units = "in")
```

@@

```{r fig.height=5, fig.width=12}
g2p_indications <- filter(read.delim("./data_munging/data_mutation_associations_appended.csv", sep = "\t", row.names = 1, header = TRUE), Evidence.Level == "A")
maf_g2p_indications <- filter(maf_raw, Genome_Change %in% g2p_indications$MutationName)
crispr_g2p_indications <- merge(maf_g2p_indications, crispr_data_ptmuts, by = c("Hugo_Symbol", "CCLE_Name", "Broad_ID"))
dup_g2p_indications <- filter(crispr_g2p_indications[, c("Hugo_Symbol", "CCLE_Name", "Broad_ID")] %>% group_by(Hugo_Symbol, CCLE_Name, Broad_ID) %>% tally(), n > 1)
crispr_g2p_indications <- merge(crispr_g2p_indications, dup_g2p_indications, by = c("Hugo_Symbol", "Broad_ID"), all.x = TRUE)
crispr_g2p_indications <- filter(crispr_g2p_indications, is.na(n))

crispr_g2p_indications_plot <- ggplot(crispr_g2p_indications, aes(x = Hugo_Symbol, y = Score)) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(alpha = 0.5, color = "lightgray", outlier.shape = NA) +
  geom_jitter(width = 0.3, mapping = aes(color = group_general_lineage_name), size = 1) +
  theme_light() +
  theme(legend.title = element_blank()) +
  labs(x = "Gene", y = "CERES Score")
crispr_g2p_indications_plot

# ggsave("./plots_18Q3/manuscript/crispr_g2p_indications_plot_test.png", crispr_g2p_indications_plot, width = 12, height = 5, units = "in")
```
